---
title: "Simple Regress"
author: "Tyler Hamblin"
date: "13 October 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
chest <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chest)
summary.lm(chest)
```


The trajectory of the CHEST data follows the expected relationship of another variable explaining LENGTH fairly well. This can be expected to perform fairly similarly to the NECK data. 


```{r}
headlen <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlen)
summary.lm(headlen)
```

The data of HEADLEN is grouped very closely and many of the data points overlap suggesting that this could be the best option for a variable determining LENGTH. 




```{r}
width <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(width)
summary.lm(width)
```

The data under this group is fairly scattered and performs the worst out of any option we have seen thus far. 


```{r}
age <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(age)
summary.lm(age)
```

This data grouping is similar to the previous one and gives no certainty as to consistently determining the LENGTH of a bear based on AGE. This will most likely be placed at the bottom of the ranking near HEADWTH. 




The list, from best to worst, of the different variables in relation to the length of a bear by their residual standard error and their adjusted R-squared:



1. LENTH~HEADLEN; Residual Standard Error: 4.235 on 52 degrees of freedom, Adjusted R-Squared: 0.8434

2. LENTH~CHEST; Residual Standard Error: 4.939 on 52 degrees of freedom, Adjusted R-Squared: 0.7869

3. LENTH~NECK; Residual Standard Error: 5.376 on 52 degrees of freedom, Adjusted R-Squared: 0.7476

4. LENTH~WEIGHT; Residual Standard Error: 5.4 on 52 degrees of freedom, Adjusted R-Squared: 0.7

5. LENTH~HEADWTH; Residual Standard Error: 7.32 on 52 degrees of freedom, Adjusted R-Squared: 0.5321

6. LENTH~AGE; Residual Standard Error: 7.5 on 52 degrees of freedom, Adjusted R-Squared: 0.50


## Conclusion

In this assignment we ran linear regression models for each of the continuous or discrete numerical variables when deciding what factor is most recognizable in correlation to the LENGTH of a bear. A part of the code chunks was included to read out the residual standard error (smaller is better) and the adjusted R-Squared values (bigger is better). According to these rules the most accurate predictor of overall length is the length of a bears head. 


