---
title: "Simple Regress"
author: "Tyler Hamblin"
date: "13 October 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
chest <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chest)
summary.lm(chest)
```



```{r}
headlen <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlen)
summary.lm(headlen)
```



```{r}
width <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(width)
summary.lm(width)
```




```{r}
age <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(age)
summary.lm(age)
```




The list, from best to worst, of the different variables in relation to the length of a bear by their residual standard error and their adjusted R-squared:



1. LENTH~HEADLEN; Residual Standard Error: 4.235 on 52 degrees of freedom, Adjusted R-Squared: 0.8434

2. LENTH~CHEST; Residual Standard Error: 4.939 on 52 degrees of freedom, Adjusted R-Squared: 0.7869

3. LENTH~NECK; Residual Standard Error: 5.376 on 52 degrees of freedom, Adjusted R-Squared: 0.7476

4. LENTH~WEIGHT; Residual Standard Error: 5.4 on 52 degrees of freedom, Adjusted R-Squared: 0.7

5. LENTH~HEADWTH; Residual Standard Error: 7.32 on 52 degrees of freedom, Adjusted R-Squared: 0.5321

6. LENTH~AGE; Residual Standard Error: 7.5 on 52 degrees of freedom, Adjusted R-Squared: 0.50


## Conclusion

In this assignment we ran linear regression models for each of the continuous or discrete numerical variables when deciding what factor is most recognizable in correlation to the LENGTH of a bear. A part of the code chunks was included to read out the residual standard error (smaller is better) and the adjusted R-Squared values (bigger is better). According to these rules the most accurate predictor of overall length is the length of a bears head. 


